Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down

### 2023-03-16 paper/ms.tex
Add google drive link

% region preamble

\documentclass[fleqn,usenatbib]{mnras}

\usepackage[T1]{fontenc}

% Allow "Thomas van Noord" and "Simon de Laguarde" and alike to be sorted by "N" and "L" etc. in the bibliography.
% Write the name in the bibliography as "\VAN{Noord}{Van}{van} Noord, Thomas"
\DeclareRobustCommand{\VAN}[3]{#2}
\let\VANthebibliography\thebibliography
\def\thebibliography{\DeclareRobustCommand{\VAN}[3]{##3}\VANthebibliography}

@@

% endregion

% region title
%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Stellar Bars and Gas]{Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down}

\author[A. Beane et al.]{Angus Beane,$^{1}$\thanks{E-mail: angus.beane@cfa.harvard.edu (AB)}
Lars Hernquist,$^{1}$
Elena~D'Onghia,$^{2,3}$
Federico~Marinacci,$^{4}$
Charlie Conroy,$^{1}$
Jia~Qi,$^{5}$\newauthor
Laura~V.~Sales,$^{6}$
Paul~Torrey,$^{5}$
Mark~Vogelsberger$^{7}$
\\
% List of institutions
$^{1}$Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA\\
$^{2}$Department of Physics, University of Wisconsin-Madison, Madison, WI, USA\\
$^{3}$Department of Astronomy, University of Wisconsin-Madison, Madison, WI, USA\\
$^{4}$Department of Physics \& Astronomy `Augusto Righi', University of Bologna, Bologna, Italy\\
$^{5}$Department of Astronomy, University of Florida, Gainesville, FL, USA\\
$^{6}$Department of Physics \& Astronomy, University of California, Riverside, CA, USA\\
$^{7}$Department of Physics, Massachusetts Institute of Technology, Cambridge, MA, USA\\
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2022}

% endregion

% Don't change these lines
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

% Abstract of the paper
\begin{abstract}
Elongated bar-like features are ubiquitous in galaxies, occurring at the centers
of approximately two-thirds of spiral disks.  Due to gravitational interactions
between the bar and the other components of galaxies, it is expected that
angular momentum and matter will redistribute over long (Gyr) timescales in
barred galaxies. Previous work ignoring the gas phase of galaxies has
conclusively demonstrated that bars should slow their rotation over time due to
their interaction with dark matter halos. We have performed a simulation of a
Milky Way-like galactic disk hosting a strong bar which includes a
state-of-the-art model of the interstellar medium and a live dark matter halo.
In this simulation the bar pattern does not slow down over time, and instead
remains at a stable, constant rate of rotation. This behavior has been observed
in previous simulations using more simplified models for the interstellar gas,
but the apparent lack of secular evolution has remained unexplained. We propose
that the gas phase of the disk and the dark matter halo act in concert to
stabilize the bar pattern speed and prevent the bar from slowing down or
speeding up. We find that in a Milky Way-like disk, a gas fraction of only about
5\% is necessary for this mechanism to operate. Our result naturally explains
why nearly all observed bars rotate rapidly and is especially relevant for our
understanding of how the Milky Way arrived at its present state.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
galaxies: bar - galaxies: disc - galaxies: evolution - galaxies: kinematics and
dynamics
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:intro}
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{fig/mockHST.pdf}
    \caption{Synthetic Hubble Space Telescope images of our simulations with and
    without the interstellar medium. The upper panels show an \Nbody{} only
    simulation while the lower panels show a simulation which includes the
    SMUGGLE model for the interstellar medium. Each panel is $20\,\textrm{kpc}$
    to a side. Columns show different points in time, separated by
    $1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar grows in
    length and strength. In the SMUGGLE run, the bar remains at approximately
    the same length and strength over the course of the simulation. The \Nbody{}
    model is identical to the GALAKOS model, discussed in the text. Details on
    the production of these synthetic images is given in Appendix~\ref{app:hst}.}\label{fig:overview}
\end{figure*}

Approximately two-thirds of spiral disks host an elongated bar-like feature at
their centers \citep{2000AJ....119..536E, 2007ApJ...657..790M}, including our
own Milky Way \citep{1957AJ.....62...19J, 1991ApJ...379..631B}. The ubiquity of
bars is not difficult to explain, since stellar disks simulated in isolation
almost always form bar-like structures \citep{1971ApJ...168..343H}. Several
studies have shown that a spherical potential, such as a stellar bulge or dark
matter halo, acts to stabilize stellar disks against bar formtaion
\citep[e.g.,][]{1973ApJ...186..467O, 1976AJ.....81...30H}.

It is more difficult to reconcile numerical simulations with the observed
pattern speeds of extragalactic bars. Currently, the best technique for
measuring the pattern speeds of individual galaxies is the Tremaine-Weinberg
method \citep{1984ApJ...282L...5T, 2011MSAIS..18...23C}. This approach has
recently been applied to samples of galaxies from the MaNGA survey
\citep{2019MNRAS.482.1733G, 2020MNRAS.491.3655G}. These studies confirm
what was found in earlier works, that nearly all extragalactic bars are fast
rotators (i.e., they rotate close to their maximum rotation rate)

This is a problem for theoretical simulations, for which there is ample evidence
that galactic bars should resonantly interact with the dark matter halo, causing
the bar to slow down over time \citep{1992ApJ...400...80H, 2000ApJ...543..704D,
2002MNRAS.330...35A, 2002ApJ...569L..83A, 2003MNRAS.341.1179A,
2003MNRAS.346..251O, 2005MNRAS.363..991H, 2006ApJ...637..214M,
2007MNRAS.375..460W, 2009ApJ...697..293D}. The physical mechanism of this
interaction can be understood as an angular form of dynamical friction between
the bar and the dark matter halo. While studied in detail for the bar
\citep{1984MNRAS.209..729T, 1985MNRAS.213..451W}, this process is generic for
any non-axisymmetric disturbance \citep{1972MNRAS.157....1L}. For an old but
still useful review of bar dynamics, see \citet{1993RPPh...56..173S}

Bar pattern speeds are usually measured using the parameter $\Rot=\RCR/\Rb$,
where \RCR{} is the corotation radius and \Rb{} is the bar length.\footnote{The
radius of corotation \RCR{} is defined for circular orbits as the radius at
which the orbital frequency is equal to the pattern speed, \PS{}, of a given
non-axisymmetric feature. In a galaxy with a constant circular velocity \Vc{},
it is given by $\RCR = \Vc / \PS$.} Galaxies with $\Rot < 1.4$ are considered
``fast rotators'' while galaxies with $\Rot > 1.4$ are considered ``slow
rotators'' \citep{2000ApJ...543..704D}. Galaxies with $\Rot < 1$ are not thought
to be stable \citep{1980AA....81..198C}. Observational estimates of the pattern
speeds of bars indicate that nearly all galaxies have $1 < \Rot <
1.4$ \citep{2011MSAIS..18...23C, 2015AA...576A.102A, 2019MNRAS.482.1733G,
2020MNRAS.491.3655G}.

@@

We have performed a simulation of a disk galaxy using the finite-volume,
gravito-hydrodynamics code \AREPO{} \citep{2010MNRAS.401..791S}. We use the galaxy
formation model Stars and MUltiphase Gas in GaLaxiEs
\citep[SMUGGLE;][]{2019MNRAS.489.4233M}. This disk galaxy exhibits almost no
evolution in the bar pattern speed over several Gyr when the gas phase is
accounted for and properly modeled. This behavior has been observed in a few
previous works \citep{1993AA...268...65F, 2010ApJ...719.1470V,
2014MNRAS.438L..81A}. However, we are the first to demonstrate that the behavior
ought to be present in a large fraction of galaxies, since it occurs in our
simulation with gas fractions as low as about $5\%$. We are also the first to
provide a physical explanation of the behavior with clear observational
predictions.

@@

We present the time evolution of different bar properties in Fig.~\ref{fig:prop}.
In the upper panel, we show the pattern speed over time in the \Nbody{} (blue)
and SMUGGLE (orange) runs. The pattern speed in the \Nbody{} case slows down
while the pattern speed in the SMUGGLE case remains roughly constant. The
slowing down of the pattern speed in the \Nbody{} case is consistent with a long
line of numerical research on bars in \Nbody{} simulations
\citep{1992ApJ...400...80H, 2000ApJ...543..704D, 2002MNRAS.330...35A,
2002ApJ...569L..83A, 2003MNRAS.341.1179A, 2003MNRAS.346..251O,
2005MNRAS.363..991H, 2006ApJ...637..214M, 2007MNRAS.375..460W,
2009ApJ...697..293D}.

However, in the SMUGGLE case the pattern speed remains constant. After the first
Gyr of evolution, we find that the pattern speed increases by only $\sim10\%$
over the next $4\,\textrm{Gyr}$, compared to a $\sim43\%$ decrease in the
pattern speed for the \Nbody{} run over the same interval.

The bottom panel of Fig.~\ref{fig:prop} shows the torque exerted on the bar by
different components. The solid lines indicate the torque on the bar by the dark
matter halo, whereas the dashed line is the torque on the bar by the gas phase.
In the \Nbody{} case, the halo exerts a steady negative torque on the bar, with
an average torque from $1$ to $4\,\textrm{Gyr}$ of $-58.0$ in units of
$10^{10}M_{\odot}\,(\textrm{km}/\textrm{s})^2$. The halo in the SMUGGLE case
exerts a similar negative torque on the bar in the first Gyr of evolution, but
after that the halo exerts a much weaker torque on the bar, averaging only
$-7.8$ in the same units and over the same time interval. The gas in the SMUGGLE
case exerts a steady positive torque averaging $11.7$ over $1\,\textrm{Gyr}$ in
the same units.

As we saw qualitatively in Fig.~\ref{fig:overview}, the upper panel of
Fig.~\ref{fig:strength} shows that the length of the bar in the \Nbody{} case
grows over time while it remains roughly constant in the SMUGGLE case. This is
also consistent with previous numerical work, which found that bars tend to grow
as they slow down and the radius of corotation increases
\citep{2000ApJ...543..704D, 2003MNRAS.341.1179A}. The time evolution of the bar
strength, defined as the maximum of $\left|A_2/A_0\right|$ as a function of
radius, is shown in the lower panel of Fig.~\ref{fig:strength}. The quantity
$\left|A_2/A_0\right|$ varies from $0$ to $1$, with larger values indicating a
stronger bar pattern. We see that in the \Nbody{} case, $\left|A_2/A_0\right|$
increases over time as the bar pattern slows. This is consistent with previous
\Nbody{} simulations which showed a clear correlation between the bar pattern
speed and the bar strength \citep[e.g.,][]{2003MNRAS.341.1179A}. In the SMUGGLE
case, we see that the bar strength has an initial drop but then remains at a
roughly constant, but slightly decreasing, strength. This is consistent with the
pattern speed in the SMUGGLE case being roughly constant or slightly increasing.

% \subsection{Figures and tables}

% Figures and tables should be placed at logical positions in the text. Don't
% worry about the exact layout, which will be handled by the publishers.

% Figures are referred to as e.g. Fig.~\ref{fig:example_figure}, and tables as
% e.g. Table~\ref{tab:example_table}.

% % Example figure
% \begin{figure}
% 	% To include a figure from a file named example.*
% 	% Allowable file formats are eps or ps if compiling using latex
% 	% or pdf, png, jpg if compiling using pdflatex
% 	\includegraphics[width=\columnwidth]{example}
%     \caption{This is an example figure. Captions appear below each figure.
% 	Give enough detail for the reader to understand what they're looking at,
% 	but leave detailed discussion to the main body of the text.}
%     \label{fig:example_figure}
% \end{figure}

% % Example table
% \begin{table}
% 	\centering
% 	\caption{This is an example table. Captions appear above each table.
% 	Remember to define the quantities, symbols and units used.}
% 	\label{tab:example_table}
% 	\begin{tabular}{lccr} % four columns, alignment for each
% 		\hline
% 		A & B & C & D\\
% 		\hline
% 		1 & 2 & 3 & 4\\
% 		2 & 4 & 6 & 8\\
% 		3 & 5 & 7 & 9\\
% 		\hline
% 	\end{tabular}
% \end{table}

>>>>

%% Beginning of file 'sample631.tex'
%%
%% Modified 2021 March
%%
%% This is a sample manuscript marked up using the
%% AASTeX v6.31 LaTeX 2e macros.
%%
%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
%% (Copyright 2000-2015).  See the classfile for details.

%% AASTeX requires revtex4-1.cls and other external packages such as
%% latexsym, graphicx, amssymb, longtable, and epsf.  Note that as of 
%% Oct 2020, APS now uses revtex4.2e for its journals but remember that 
%% AASTeX v6+ still uses v4.1. All of these external packages should 
%% already be present in the modern TeX distributions but not always.
%% For example, revtex4.1 seems to be missing in the linux version of
%% TexLive 2020. One should be able to get all packages from www.ctan.org.
%% In particular, revtex v4.1 can be found at 
%% https://www.ctan.org/pkg/revtex4-1.

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%% using aastex version 6.3
\documentclass[twocolumn,linenumbers,trackchanges]{aastex631}

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be invoked like this:
%%
%% \documentclass[arguments]{aastex631}
%% 
%% where the layout options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.  
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%  RNAAS       : Supresses an abstract. Originally for RNAAS manuscripts 
%%                but now that abstracts are required this is obsolete for
%%                AAS Journals. Authors might need it for other reasons. DO NOT
%%                use \begin{abstract} and \end{abstract} with this style.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can invoke to allow other stylistic
%% actions. The available options are:
%%
%%   astrosymb    : Loads Astrosymb font and define \astrocommands. 
%%   tighten      : Makes baselineskip slightly smaller, only works with 
%%                  the twocolumn substyle.
%%   times        : uses times font instead of the default
%%   linenumbers  : turn on lineno package.
%%   trackchanges : required to see the revision mark up and print its output
%%   longauthor   : Do not use the more compressed footnote style (default) for 
%%                  the author/collaboration/affiliations. Instead print all
%%                  affiliation information after each name. Creates a much 
%%                  longer author list but may be desirable for short 
%%                  author papers.
%% twocolappendix : make 2 column appendix.
%%   anonymous    : Do not show the authors, affiliations and acknowledgments 
%%                  for dual anonymous review.
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex631}
%%
%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%% \hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
%%
%% Note that in v6.3 "bookmarks" has been changed to "true" in hyperref
%% to improve the accessibility of the compiled pdf file.
%%
%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%

@@

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
%\received{March 1, 2021}
%\revised{April 1, 2021}
%\accepted{\today}

\shorttitle{Stellar Bars and Gas}
\shortauthors{Beane et al.}

\graphicspath{{./}{figures/}}

\begin{document}
\title{Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire. In v6.31 you can include a footnote in the title.

%% A significant change from earlier AASTEX versions is in the structure for 
%% calling author and affiliations. The change was necessary to implement 
%% auto-indexing of affiliations which prior was a manual process that could 
%% easily be tedious in large author manuscripts.
%%
%% The \author command is the same as before except it now takes an optional
%% argument which is the 16 digit ORCID. The syntax is:
%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
%%
%% This will hyperlink the author name to the author's ORCID page. Note that
%% during compilation, LaTeX will do some limited checking of the format of
%% the ID to make sure it is valid. If the "orcid-ID.png" image file is 
%% present or in the LaTeX pathway, the OrcID icon will appear next to
%% the authors name.
%%
%% Use \affiliation for affiliation information. The old \affil is now aliased
%% to \affiliation. AASTeX v6.31 will automatically index these in the header.
%% When a duplicate is found its index will be the same as its previous entry.
%%
%% Note that \altaffilmark and \altaffiltext have been removed and thus 
%% can not be used to document secondary affiliations. If they are used latex
%% will issue a specific error message and quit. Please use multiple 
%% \affiliation calls for to document more than one affiliation.
%%
%% The new \altaffiliation can be used to indicate some secondary information
%% such as fellowships. This command produces a non-numeric footnote that is
%% set away from the numeric \affiliation footnotes.  NOTE that if an
%% \altaffiliation command is used it must come BEFORE the \affiliation call,
%% right after the \author command, in order to place the footnotes in
%% the proper location.
%%
%% Use \email to set provide email addresses. Each \email will appear on its
%% own line so you can put multiple email address in one \email call. A new
%% \correspondingauthor command is available in V6.31 to identify the
%% corresponding author of the manuscript. It is the author's responsibility
%% to make sure this name is also in the author list.
%%
%% While authors can be grouped inside the same \author and \affiliation
%% commands it is better to have a single author for each. This allows for
%% one to exploit all the new benefits and should make book-keeping easier.
%%
%% If done correctly the peer review system will be able to
%% automatically put the author and affiliation information from the manuscript
%% and save the corresponding author the trouble of entering it by hand.

%\correspondingauthor{August Muench}
%\email{greg.schwarz@aas.org, gus.muench@aas.org}

\author{Angus Beane}
\affiliation{Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA}

\author{Lars Hernquist}
\affiliation{Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA}

\author{Elena D'Onghia}
\affiliation{Department of Physics, University of Wisconsin-Madison, Madison, WI, USA}
\affiliation{Department of Astronomy, University of Wisconsin-Madison, Madison, WI, USA}

\author{Federico Marinacci}
\affiliation{Department of Physics \& Astronomy `Augusto Righi', University of Bologna, Bologna, Italy}

\author{Charlie Conroy}
\affiliation{Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA}

\author{Jia Qi}
\affiliation{Department of Astronomy, University of Florida, Gainesville, FL, USA}

\author{Laura V. Sales}
\affiliation{Department of Physics \& Astronomy, University of California, Riverside, CA, USA}

\author{Paul Torrey}
\affiliation{Department of Astronomy, University of Florida, Gainesville, FL, USA}

\author{Mark Vogelsberger}
\affiliation{Department of Physics, Massachusetts Institute of Technology, Cambridge, MA, USA}

\begin{abstract}

  Elongated bar-like features are ubiquitous in galaxies, occurring at the
  centers of approximately two-thirds of spiral disks in the nearby Universe.  
  Due to gravitational
  interactions between the bar and the other components of galaxies, it is
  expected that angular momentum and matter will redistribute over long (Gyr)
  timescales in barred galaxies. Previous work ignoring the gas phase of
  galaxies has conclusively demonstrated that bars should slow their rotation
  over time due to their interaction with dark matter halos. We have performed a
  simulation of a Milky Way-like galactic disk hosting a strong bar which
  includes a state-of-the-art model of the interstellar medium and a live dark
  matter halo. In this simulation the bar pattern does not slow down over time,
  and instead remains at a stable, constant rate of rotation. This behavior has
  been observed in previous simulations using more simplified models for the
  interstellar gas, but the apparent lack of secular evolution has remained
  unexplained. We propose that the gas phase of the disk and the dark matter
  halo act in concert to stabilize the bar pattern speed and prevent the bar
  from slowing down or speeding up. We find that in a Milky Way-like disk, a gas
  fraction of only about 5\% is necessary for this mechanism to operate. Our
  result naturally explains why nearly all observed bars rotate rapidly and is
  especially relevant for our understanding of how the Milky Way arrived at its
  present state.

\end{abstract}

%% Keywords should appear after the \end{abstract} command. 
%% The AAS Journals now uses Unified Astronomy Thesaurus concepts:
%% https://astrothesaurus.org
%% You will be asked to selected these concepts during the submission process
%% but this old "keyword" functionality is maintained in case authors want
%% to include these concepts in their preprints.
\keywords{Classical Novae (251) --- Ultraviolet astronomy(1736) --- History of astronomy(1868) --- Interdisciplinary astronomy(804)}

%% From the front matter, we move on to the body of the paper.
%% Sections are demarcated by \section and \subsection, respectively.
%% Observe the use of the LaTeX \label
%% command after the \subsection to give a symbolic KEY to the
%% subsection for cross-referencing in a \ref command.
%% You can use LaTeX's \ref and \label commands to keep track of
%% cross-references to sections, equations, tables, and figures.
%% That way, if you change the order of any elements, LaTeX will
%% automatically renumber them.
%%
%% We recommend that authors also use the natbib \citep
%% and \citet commands to identify citations.  The citations are
%% tied to the reference list via symbolic KEYs. The KEY corresponds
%% to the KEY in the \bibitem in the reference list below.

\section{Introduction}
\label{sec:intro}
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{fig/mockHST.pdf}
    \caption{Synthetic Hubble Space Telescope images of our simulations with and
    without the interstellar medium. The upper panels show an \Nbody{} only
    simulation while the lower panels show a simulation which includes the
    \SMUGGLE{} model for the interstellar medium. Each panel is $20\,\textrm{kpc}$
    to a side. Columns show different points in time, separated by
    $1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar grows in
    length and strength. In the \SMUGGLE{} run, the bar remains at approximately
    the same length and strength over the course of the simulation. The \Nbody{}
    model is identical to the GALAKOS model, discussed in the text. Details on
    the production of these synthetic images is given in Appendix~\ref{app:hst}.}\label{fig:overview}
\end{figure*}

Approximately two-thirds of spiral disks host an elongated bar-like feature at
their centers \citep{2000AJ....119..536E, 2007ApJ...657..790M}, including our
own Milky Way \citep{1957AJ.....62...19J, 1991ApJ...379..631B}. The ubiquity of
bars is not difficult to explain, since stellar disks simulated in isolation
almost always form bar-like structures \citep{1971ApJ...168..343H}. Several
studies have shown that a spherical potential, such as a stellar bulge or dark
matter halo, acts to stabilize stellar disks against bar formation
\citep[e.g.,][]{1973ApJ...186..467O, 1976AJ.....81...30H}.

It is more difficult to reconcile numerical simulations with the observed
pattern speeds of extragalactic bars. Currently, the best technique for
measuring the pattern speeds of individual galaxies is the Tremaine-Weinberg
method \citep{1984ApJ...282L...5T, 2011MSAIS..18...23C}. This approach has
recently been applied to samples of galaxies from the MaNGA survey
\citep{2019MNRAS.482.1733G, 2020MNRAS.491.3655G}. These studies confirm
what was found in earlier works, that nearly all extragalactic bars are fast
rotators (i.e., they rotate close to their maximum rotation rate).

This is a problem for theoretical simulations, for which there is ample evidence
that galactic bars should resonantly interact with the dark matter halo, causing
the bar to slow down over time \citep{1992ApJ...400...80H, 2000ApJ...543..704D,
2002MNRAS.330...35A, 2002ApJ...569L..83A, 2003MNRAS.341.1179A,
2003MNRAS.346..251O, 2005MNRAS.363..991H, 2006ApJ...637..214M,
2007MNRAS.375..460W, 2009ApJ...697..293D}. The physical mechanism of this
interaction can be understood as an angular form of dynamical friction between
the bar and the dark matter halo. While studied in detail for the bar
\citep{1984MNRAS.209..729T, 1985MNRAS.213..451W}, this process is generic for
any non-axisymmetric disturbance \citep{1972MNRAS.157....1L}. (For an old but
still useful review of bar dynamics, see \citet{1993RPPh...56..173S}.)

Bar pattern speeds are usually measured using the parameter $\Rot=\RCR/\Rb$,
where \RCR{} is the corotation radius and \Rb{} is the bar length.\footnote{The
radius of corotation \RCR{} is defined for circular orbits as the radius at
which the orbital frequency is equal to the pattern speed, \PS{}, of a given
non-axisymmetric feature. In a galaxy with a constant circular velocity \Vc{},
it is given by $\RCR = \Vc / \PS$.} Galaxies with $\Rot < 1.4$ are considered
``fast rotators'' while galaxies with $\Rot > 1.4$ are considered ``slow
rotators'' \citep{2000ApJ...543..704D}. Galaxies with $\Rot < 1$ are not thought
to be stable \citep{1980AA....81..198C}. Observational estimates of the pattern
speeds of bars indicate that nearly all galaxies have $1 < \Rot < 1.4$
\citep{2011MSAIS..18...23C, 2015AA...576A.102A, 2019MNRAS.482.1733G,
2020MNRAS.491.3655G}. We note that \citet{2017ApJ...835..279F} argue that the
pattern speed should be measured relative to a characteristic angular velocity
of the outer disk.

@@

In the last decade, stellar bars have been studied mainly in the
context of the instability processes that lead to their formation and their
ability to drive gas toward the galaxy center and contribute to the formation of
supermassive black holes (SMBH). The primary interest of these studies mainly
was in the loss of angular momentum of the gas and the associated  gaseous flow
down to the inner disk, possibly forming a central mass concentration
\citep{2010ApJ...719.1470V}, and fueling the central SMBH \citep[e.g.][]{1989Natur.338...45S, 1990Natur.345..679S}. A revisit of the more general problem of
galactic bar properties and formation, including the case of disk galaxies with
very large gas fraction is timely since it is clear now that galactic disks show
massive bars already at redshift $z>2$ \citep{2022arXiv221008658G}. At
that time the universe was 2.5 billion year old and galaxies might have as high
gas fraction as 80\% \citep{2020ARAA..58..157T}. Furthermore, unlike
nearby disk galaxies, the high-redshift disks also continuously accrete cold gas
from the cosmic web, making the formation, stability and evolution of
non-axisymmetric features a key question to address since they can play a
fundamental role in the more general problem of disk galaxy evolution.

We have performed a simulation of a disk galaxy using the finite-volume,
gravito-hydrodynamics code \AREPO{} \citep{2010MNRAS.401..791S}. We use the
galaxy formation model Stars and MUltiphase Gas in GaLaxiEs
\citep[\SMUGGLE{};][]{2019MNRAS.489.4233M}. This disk galaxy exhibits almost no
evolution in the bar pattern speed over several Gyr when the gas phase is
accounted for and properly modeled. This behavior has been observed in a few
previous works \citep{1993AA...268...65F, 2007ApJ...666..189B,
2009ApJ...707..218V, 2010ApJ...719.1470V, 2014MNRAS.438L..81A}.

Here we show that this phenomenon occurs in galaxies with gas fraction
as low as about $5\%$, making this feature a likely characteristic of  Milky
Way-type galaxies. This behavior will be explored in more general disks with
larger gas fractions than the Milky Way in a forthcoming paper.
% We are also the first to provide a physical explanation of
%the behavior with clear observational predictions.
We also provide a physical explanation of the redistribution of the angular
momentum among the dark halo wake, the stellar bar, and the disk gas component.

@@

We present the time evolution of different bar properties in Fig.~\ref{fig:prop}.
In the upper panel, we show the pattern speed over time in the \Nbody{} (blue)
and \SMUGGLE{} (orange) runs. The pattern speed in the \Nbody{} case slows down
while the pattern speed in the \SMUGGLE{} case remains roughly constant. The
slowing down of the pattern speed in the \Nbody{} case is consistent with a long
line of numerical research on bars in \Nbody{} simulations
\citep{1992ApJ...400...80H, 2000ApJ...543..704D, 2002MNRAS.330...35A,
2002ApJ...569L..83A, 2003MNRAS.341.1179A, 2003MNRAS.346..251O,
2005MNRAS.363..991H, 2006ApJ...637..214M, 2007MNRAS.375..460W,
2009ApJ...697..293D}.

However, in the \SMUGGLE{} case the pattern speed remains constant. After the first
Gyr of evolution, we find that the pattern speed increases by only $\sim10\%$
over the next $4\,\textrm{Gyr}$, compared to a $\sim43\%$ decrease in the
pattern speed for the \Nbody{} run over the same interval.

The bottom panel of Fig.~\ref{fig:prop} shows the torque exerted on the bar by
different components. The solid lines indicate the torque on the bar by the dark
matter halo, whereas the dashed line is the torque on the bar by the gas phase.
In the \Nbody{} case, the halo exerts a steady negative torque on the bar, with
an average torque from $1$ to $4\,\textrm{Gyr}$ of $-58.0$ in units of
$10^{10}M_{\odot}\,(\textrm{km}/\textrm{s})^2$. The halo in the \SMUGGLE{} case
exerts a similar negative torque on the bar in the first Gyr of evolution, but
after that the halo exerts a much weaker torque on the bar, averaging only
$-7.8$ in the same units and over the same time interval. The gas in the \SMUGGLE{}
case exerts a steady positive torque averaging $11.7$ over $1\,\textrm{Gyr}$ in
the same units.

As we saw qualitatively in Fig.~\ref{fig:overview}, the upper panel of
Fig.~\ref{fig:strength} shows that the length of the bar in the \Nbody{} case
grows over time while it remains roughly constant in the \SMUGGLE{} case. This is
also consistent with previous numerical work, which found that bars tend to grow
as they slow down and the radius of corotation increases
\citep{2000ApJ...543..704D, 2003MNRAS.341.1179A}. The time evolution of the bar
strength, defined as the maximum of $\left|A_2/A_0\right|$ as a function of
radius, is shown in the lower panel of Fig.~\ref{fig:strength}. The quantity
$\left|A_2/A_0\right|$ varies from $0$ to $1$, with larger values indicating a
stronger bar pattern. We see that in the \Nbody{} case, $\left|A_2/A_0\right|$
increases over time as the bar pattern slows. This is consistent with previous
\Nbody{} simulations which showed a clear correlation between the bar pattern
speed and the bar strength \citep[e.g.,][]{2003MNRAS.341.1179A}. In the \SMUGGLE{}
case, we see that the bar strength has an initial drop but then remains at a
roughly constant, but slightly decreasing, strength. This is consistent with the
pattern speed in the \SMUGGLE{} case being roughly constant or slightly increasing.

\section{Discussion}
\label{sec:discussion}
\subsection{Pattern Speed Evolution}
The lack of evolution in the pattern speed of the \SMUGGLE{} case (seen in
Fig.~\ref{fig:prop}) is intimately tied to the sudden decrease in torque exerted
on the bar by the dark matter halo.

%We argue that this can be understood in terms of the halo wake mechanism. 
We interpret this behavior in terms of the halo wake mechanism. In the
\Nbody{} case, a well known phenomenon is that the halo material resonant with
the bar forms a wake, and this wake lags behind \citep{1984MNRAS.209..729T,
1985MNRAS.213..451W, 1992ApJ...400...80H} and exerts a negative torque on the
bar, slowing it down (see Fig.~\ref{fig:wake} below).\footnote{Since the bar is not a solid
body, it is not guaranteed that a negative torque will slow it down - e.g. a
negative torque could shred the bar, reducing its moment of inertia without
changing its pattern speed. However, the bar seems to empirically respond to a
negative torque induced by a halo wake by slowing down.} As the bar slows down,
the location of the resonances in the phase space changes (see Fig.~12 and Table
1 in \citet{2020ApJ...890..117D}) allowing halo material newly resonant with the
bar to participate in the formation of the wake. However, the gas is also a
reliable source of positive torque on the bar, speeding the bar up. In turn,
this stops the location of the resonance from changing such that the halo cannot
reinforce the wake, therefore arresting the process by which the halo can slow
the bar down. We term this process ``halo exhaustion.'' Note that the resonances
can be maintained even in the presence of the dissipative gaseous component if
the gas is continuously accreted from the external environment, as is the case
for massive disk galaxies at high redshift.

%In the \Nbody{} case, halo material that is
%resonant with the bar will form a wake which lags behind and exerts a negative 
%torque on the bar, slowing it down \citep{1984MNRAS.209..729T,
%1985MNRAS.213..451W, 1992ApJ...400...80H}.\footnote{Since the bar is not a solid
%body, it is not guaranteed that a negative torque will slow it down - e.g. a
%negative torque could shred the bar, reducing its moment of inertia without
%changing its pattern speed. However, the bar seems to empirically respond to a
%negative torque induced by a halo wake by slowing down.} As the bar slows down,
%the location of the resonances in phase space changes, allowing halo material
%newly resonant with the bar to participate in forming the wake. However, the gas
%is a reliable source of positive torque on the bar, speeding the bar up. This
%stops the resonance location from changing such that the halo cannot reinforce
%the wake, arresting the process by which the halo can slow the bar down. We term
%this process ``halo exhaustion.''

This halo exhaustion process is similar to the ``metastability'' effect, which
has been previously discussed in the literature \citep{2003MNRAS.345..406V,
2006ApJ...639..868S}. Finally, we also note that these authors, in particular,
observe that the effects of numerical resolution in the simulations adopted to
explore these mechanisms have yet to be fully explored and could play a role in
the observed phenomenology of the simulations. We plan to address these issues
in future dedicated work.

We test this interpretation by measuring the angle offset between the halo
wake and the bar. If the wake and the bar are aligned (i.e., there is no angle
offset), then the wake cannot exert a negative torque on the bar. This angle
is plotted in the left panel of Fig.~\ref{fig:wake}, which shows that the
angle offset is larger in the \Nbody{} case than in the \SMUGGLE{} case by about a
factor of two. The center and right panels of Fig.~\ref{fig:wake} show the halo
wake with respect to the location of the bar in the \Nbody{} (center) and \SMUGGLE{}
(right) cases at one point in time.
